---
title: "Additional Parameters for the Genesis API Call"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{vignette-additional_parameter}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---




```{r eval=FALSE, echo=FALSE}
library(restatis)
```

## Information

This is a brief overview of the additional parameters for the Genesis API call. These parameters only affect the Genesis API call itself, no further processing. 

Not all parameters are available for all functions. 

Especially `gen_table()` and `gen_cube()` have a lot of different additional parameters to specify the API call. 

For a detailed overview, see the following official document: https://www-genesis.destatis.de/genesis/misc/GENESIS-Webservices_Einfuehrung.pdf

## Short Overview of the Most Used Additional Parameters

### Parameter: Selection & Searchcriterion

The `selection`-parameter is a string that specifies the filter to use for filtering the API call. The criterion to which this parameter is applied is defined by the `searchcriterion`-parameter. The maximum length of the string depends on the function being used, but often varies between 10 and 15 characters. It is possible to use '*' as a placeholder. The default for this parameter is NO filtering. 

The `searchcriterion`-parameter is a string that specifies the criteria to be used by the `selection`-parameter. Possible values depend on the function being used, but are often "code" for filtering based on code, or "content" for filtering based on content description. The default for this parameter is NO filter criterion. 

Examples:
```{r eval=FALSE}
# Get the values of the variable "WAM8" which code starts with "WA29"
gen_val2var("WAM8", selection = "WA29*", searchcriterion = "code")

# Same result can be achieved by the following line due to the default of the "searchcriterion"-parameter:
gen_val2var("WAM8", selection = "WA29*")
```

### Parameter: Sortcriterion
The `sortcriterion`-parameter is a string that specifies whether the raw output of the API call is sorted according to the specified input. Possible values depend on the function being used, but often "code" for sorting based on code and "content" for sorting based on content description are possible. The default is NO sorting.

Examples:
```{r eval=FALSE}
# Get the values of the variable "WAM8" sorted based on their codes
gen_val2var("WAM8", sortcriterion = "code")
```

### Parameter: Pagelength
The `pagelength`-parameter is a number between 1 and 2500. It indicates the maximum length of the API call output. Default is 100.

Examples:
```{r eval=FALSE}
# Get the values of the variable "WAM8", where the maximum length of the output is 2500 entries
gen_val2var("WAM8", pagelength = 2500)
```

### Parameter: Language
The `language`-parameter is a string. It indicates if the API call output is in German by using "de" or in English by using "en". Default is "de".

Examples:
```{r eval=FALSE}
# Get the values of the variable "WAM8" in German
gen_val2var("WAM8", language = "de")

# Get the values of the variable "WAM8" in English
gen_val2var("WAM8", language = "en")
```


## Explicit Overview of Additional Parameters per Function

<details>
<summary>gen_alternative_terms</summary> 

| Parameter   | Description |
| ----------- | ----------- |
| selection      | 1-15 characters; filtering based on "code" of the objects; "*"-Notations are possible       |
| pagelength   | 1-2500        |
| language   | "de" / "en"        |

</details>

<details>
<summary>gen_catalogue</summary> 
For cubes:

| Parameter   | Description |
| ----------- | ----------- |
| selection      | 1-10 characters; filtering based on "code" of the objects; "*"-Notations are possible       |
| pagelength   | 1-2500        |
| language   | "de" / "en"        |
| area   | -       |

For statistics:

| Parameter   | Description |
| ----------- | ----------- |
| selection      | 1-15 characters; filtering based on searchcriterion; "*"-Notations are possible       |
| searchcriterion   | "code" / "content"       |
| sortcriterion   | "code" / "content"       |
| pagelength   | 1-2500        |
| language   | "de" / "en"        |

For tables:

| Parameter   | Description |
| ----------- | ----------- |
| selection      | 1-15 characters; filtering based on searchcriterion; "*"-Notations are possible       |
| searchcriterion   | "code"     |
| sortcriterion   | "code" / "top"       |
| pagelength   | 1-2500        |
| language   | "de" / "en"        |
| area   | -        |

</details>

<details>
<summary>gen_cube</summary> 
see the following official documentation: https://www-genesis.destatis.de/genesis/misc/GENESIS-Webservices_Einfuehrung.pdf

</details>

<details>
<summary>gen_find</summary> 
| Parameter   | Description |
| ----------- | ----------- |
| pagelength   | 1-2500        |
| language   | "de" / "en"        |

</details>

<details>
<summary>gen_list_jobs</summary> 
| Parameter   | Description |
| ----------- | ----------- |
| selection      | 1-50 characters; filtering based on "code" of the objects or searchcriterion; "*"-Notations are possible       |
| searchcriterion   | "content" / "time" / "status"     |
| sortcriterion   | "content" / "time" / "status"      |
| pagelength   | 1-2500        |
| language   | "de" / "en"        |
| type   | -        |

see more in the following official documentation: https://www-genesis.destatis.de/genesis/misc/GENESIS-Webservices_Einfuehrung.pdf 

</details>

<details>
<summary>gen_metadata</summary> 
| Parameter   | Description |
| ----------- | ----------- |
| language   | "de" / "en"        |
| area   | -        |
</details>

<details>
<summary>gen_modified_data</summary> 
| Parameter   | Description |
| ----------- | ----------- |
| selection   | 1-15 characters; filtering based on "code" of the objects; "*"-Notations are possible |
| pagelength   | 1-2500        |
| language   | "de" / "en"        |
</details>

<details>
<summary>gen_objects2stat</summary> 
For cubes:

| Parameter   | Description |
| ----------- | ----------- |
| selection      | 1-10 characters; filtering based on "code" of the objects; "*"-Notations are possible       |
| pagelength   | 1-2500        |
| language   | "de" / "en"        |

For variables:

| Parameter   | Description |
| ----------- | ----------- |
| selection      | 1-6 characters; filtering based on searchcriterion; "*"-Notations are possible       |
| searchcriterion   | "code" / "content"       |
| sortcriterion   | "code" / "content"       |
| pagelength   | 1-2500        |
| language   | "de" / "en"        |
| area   | -        |
| type   | -        |

see more in the following official documentation: https://www-genesis.destatis.de/genesis/misc/GENESIS-Webservices_Einfuehrung.pdf 

For tables:

| Parameter   | Description |
| ----------- | ----------- |
| selection      | 1-15 characters; filtering based on "code" of the objects; "*"-Notations are possible       |
| pagelength   | 1-2500        |
| language   | "de" / "en"        |
| area   | -        |

</details>

<details>
<summary>gen_objects2var</summary> 
For cubes:

| Parameter   | Description |
| ----------- | ----------- |
| selection      | 1-15 characters; filtering based on "code" of the objects; "*"-Notations are possible       |
| pagelength   | 1-2500        |
| language   | "de" / "en"        |
| area   | -       |

For statistics:

| Parameter   | Description |
| ----------- | ----------- |
| selection      | 1-15 characters; filtering based on searchcriterion; "*"-Notations are possible       |
| searchcriterion   | "code" / "content"       |
| sortcriterion   | "code" / "content"       |
| pagelength   | 1-2500        |
| language   | "de" / "en"        |
| area   | -       |

For tables:

| Parameter   | Description |
| ----------- | ----------- |
| selection      | 1-15 characters; filtering based on "code" of the objects; "*"-Notations are possible       |
| pagelength   | 1-2500        |
| language   | "de" / "en"        |
| area   | -        |
</details>

<details>
<summary>gen_table</summary> 
see the following official documentation: https://www-genesis.destatis.de/genesis/misc/GENESIS-Webservices_Einfuehrung.pdf

</details>

<details>
<summary>gen_var2stat</summary> 
| Parameter   | Description |
| ----------- | ----------- |
| selection      | 1-6 characters; filtering based on searchcriterion; "*"-Notations are possible       |
| searchcriterion   | "code" / "content"       |
| pagelength   | 1-2500        |
| language   | "de" / "en"        |
| area   | -        |
| type   | -        |

see more in the following official documentation: https://www-genesis.destatis.de/genesis/misc/GENESIS-Webservices_Einfuehrung.pdf 

</details>

<details>
<summary>gen_val2var</summary> 
| Parameter   | Description |
| ----------- | ----------- |
| selection      | 1-15 characters; filtering based on searchcriterion; "*"-Notations are possible       |
| searchcriterion   | "code" / "content"       |
| pagelength   | 1-2500        |
| language   | "de" / "en"        |

</details>

<details>
<summary>gen_val2var2stat</summary> 
| Parameter   | Description |
| ----------- | ----------- |
| selection      | 1-6 characters; filtering based on searchcriterion; "*"-Notations are possible       |
| searchcriterion   | "code" / "content"       |
| pagelength   | 1-2500        |
| language   | "de" / "en"        |
| area   | -        |
| type   | -        |

see more in the following official documentation: https://www-genesis.destatis.de/genesis/misc/GENESIS-Webservices_Einfuehrung.pdf 

</details>

<details>
<summary>gen_search_vars</summary> 
| Parameter   | Description |
| ----------- | ----------- |
| selection      | 1-6 characters; filtering based on searchcriterion; "*"-Notations are possible       |
| searchcriterion   | "code" / "content"       |
| pagelength   | 1-2500        |
| language   | "de" / "en"        |
| area   | -        |
| type   | -        |

see more in the following official documentation: https://www-genesis.destatis.de/genesis/misc/GENESIS-Webservices_Einfuehrung.pdf 

</details>
