---
title: "Basic `{restatis}` workflow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Basic `{restatis}` workflow}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---

```{r}
library(restatis)
library(ggplot2)
library(dplyr, warn.conflicts = FALSE)
```


## Introduction

This vignette describes the basic workflow for finding, exploring, and retrieving data from German official statistics databases using the `{restatis}` package. We will walk through the entire data discovery and retrieval pipeline, from initial search to visualisation and analysis.

`{restatis}` provides access to several official German statistical databases through R, including:

- GENESIS Online: The main database of the German Federal Statistical Office (Destatis)
- regionalstatistik.de: Regional statistics data
- Zensus 2022: German Census 2022 data
- And several state-level or special-interest statistical databases

For more details on additional parameters for API calls, see `vignette("additional_parameter")`.

## Real-World Scenario: Analysing Consumer Price Trends

Throughout this vignette, we will use a realistic example: Analysing how consumer prices for specific product categories have changed since 2020. This is a common use case for economists, policy analysts, and journalists working with German official statistics.

### Step 1: Exploring Available Data

Before requesting data, it is important to understand what is available. Let us start by searching for data related to consumer prices.

#### Finding Alternative Terms

The GENESIS database is quite comprehensive, so starting with the right keywords is crucial. We can use `gen_alternative_terms()` to discover related terminology:

```{r}

# Search for terms related to consumer prices (German: "Verbraucherpreise")
restatis::gen_alternative_terms(term = "*verbraucherpr*", 
                                database = "genesis")

```

This helps us understand how the database structures price-related data. We discover that "Verbraucherpreisindex" (Consumer Price Index) is the primary term, and we can also find related concepts like specific commodity categories.

### Step 2: Finding Relevant Data Objects

Once we have our keywords, we search for actual data tables or cubes, the main data structures of GENESIS-based databases. The `gen_find()` function returns results from different categories (set `category` to `all`):

```{r}

# Search for consumer price data
search_results <- restatis::gen_find(term = "verbraucherpreisindex",
                                     detailed = FALSE,
                                     ordering = TRUE,
                                     category = "all",
                                     database = "genesis")

# View available tables
search_results$Tables

```

The function returns results organised by object type (statistics, tables, cubes, variables). Each result includes:

- `Code`: The unique identifier for the data object
- `Title`: What the data represents
- `Content`: Additional description

For our consumer price analysis, we are interested in tables that contain monthly or regular updates. Table `61111-0004` is a good candidate: "Verbraucherpreisindex: Deutschland, Monate, Klassifikation des Individualkonsums" (Consumer Price Index: Germany, Months, Personal Consumption Classification).

### Step 3: Exploring Table Metadata

Before downloading data, always check the metadata to understand the structure and available filters:

```{r}

# Get metadata for table 61111-0004
price_meta <- restatis::gen_metadata_table(code = "61111-0004",
                                           database = "genesis")

# Examine the structure
str(price_meta)

# Look at available variables (rows/columns you can filter by)
price_meta$Structure$Rows

```

The metadata reveals important information:

- Available time periods: What years/months have data?
- Classification variables: What dimensions can we filter by (e.g., product categories)?
- Value units: The units of measurement
- Variable attributes: Additional metadata for filtering

For the consumer price index, you will typically find:

- A "Sachklassifikation" (commodity classification) variable (e.g., "CC13A5") that lets you select specific product categories
- Monthly time data
- Values indexed to a base year (e.g., 2020=100)

### Step 4: Downloading Filtered Data

Now we move from exploration to data retrieval. This is where `{restatis}` becomes powerful — you can filter data at the API level, rather than downloading everything and filtering locally.

#### Example 1: Simple Time Series (No Filtering)

For a quick overview, let us first get unfiltered price index data:

```{r}

library(ggplot2)

# Download price index data for 2020 onwards
price_index <- restatis::gen_table(name = "61111-0001",
                                   language = "de",
                                   startyear = 2020,
                                   database = "genesis") %>%
  mutate(value = as.numeric(gsub(",", ".", value))) %>%
  filter(value_unit == "2020=100")

# Quick visualisation
price_index %>%
  ggplot(aes(x = time, y = value, group = 1)) +
  geom_line(linewidth = 1) +
  labs(title = "Consumer Price Index Germany",
       subtitle = "Base: 2020=100",
       x = "Year",
       y = "Index") +
  theme_minimal()

```

This shows the overall consumer price trend, but we can do much better by filtering specific product categories.

#### Example 2: Filtered Data by Product Category

The real power of `{restatis}` comes from filtering at the API level. Let us get data for specific product categories:

```{r}

# First, get metadata to understand the classification variable codes
price_meta <- restatis::gen_metadata_table(code = "61111-0004",
                                           database = "genesis")

# The classification variable is typically stored in Structure$Rows
# Look for the "Sachklassifikation" or similar classification variable
# This will tell us the variable code (e.g., "CC13A5") and available category codes

# Now retrieve filtered data for specific product categories
# Example: Food (CC13-04521), Transport (CC13-04522), Housing (CC13-01151), Clothing (CC13-02201)

price_filtered <- restatis::gen_table(name = "61111-0004",
                                      language = "de",
                                      classifyingvariable1 = "CC13A5",  # The classification variable
                                      classifyingkey1 = "CC13-04521,CC13-04522,CC13-01151,CC13-02201",  # Specific categories
                                      startyear = 2020,
                                      database = "genesis") %>%
  mutate(value = as.numeric(gsub(",", ".", value))) # Correct for the decimal sign

# Parse the time information to create proper dates
# The package returns time as strings that need parsing
# E.g., extract month from attribute code
price_filtered <- price_filtered %>%
  mutate(year_month = as.Date(paste0(time, 
                                     "-", 
                                     substr(`1_variable_attribute_code`, 6, 8), 
                                     "-15")))

# Visualisation with multiple categories
price_filtered %>%
  ggplot(aes(x = year_month, y = value, color = `3_variable_attribute_label`)) +
  geom_line(linewidth = 1) +
  scale_color_discrete(name = NULL,
                       labels = function(x) gsub(",.*", "", x)) +
  labs(title = "Consumer Price Index For Selected Product Categories",
       subtitle = "Germany, Base: 2020=100",
       x = "",
       y = "Index") +
  theme_minimal() +
  theme(legend.position = "bottom")

```

Key insight: By using the `classifyingvariable1` and `classifyingkey1` parameters, we filter data at the API level. This is more efficient than downloading all data and filtering in R.

### Step 5: Practical Data Wrangling

Once you have the data, you typically need to clean and prepare it:

```{r}

# Clean and prepare for analysis
price_data <- price_filtered %>%
  select(year_month, value, `3_variable_attribute_label`) %>%
  rename(date = year_month,
         price_index = value,
         category = `3_variable_attribute_label`) %>%
  mutate(category = gsub(",.*", "", category),  # Remove extra text after comma
         price_index = as.numeric(price_index)) %>%
  arrange(date, category)

# Calculate year-over-year changes
price_data <- price_data %>%
  group_by(category) %>%
  arrange(date) %>%
  mutate(yoy_change = (price_index - lag(price_index, 12)) / lag(price_index, 12) * 100) %>%
  ungroup()

# Filter to recent data and calculate latest changes
latest_changes <- price_data %>%
  filter(!is.na(yoy_change)) %>%
  group_by(category) %>%
  slice_tail(n = 1) %>%
  arrange(desc(yoy_change))

print(latest_changes)

```

### Step 6: Advanced Features

#### Checking for Data Updates

After collecting data, you might want to check if it has been updated:

```{r}

restatis::gen_modified_data(code = "61111",  # Prefix of the table code
                            date = "month_before",
                            database = "genesis")

```

This is useful if you are collecting data regularly and want to know what is new.

#### Using Multiple Time Periods

You can specify time ranges more precisely:

```{r}

# Get data for a specific range of years
price_range <- restatis::gen_table(name = "61111-0004",
                                   language = "de",
                                   classifyingvariable1 = "CC13A5",
                                   classifyingkey1 = "CC13-04521",  # Just food
                                   startyear = 2015,
                                   endyear = 2025,
                                   database = "genesis")

```

#### Important Notes on Data Handling

- Decimal separators: German databases use commas for decimals (note: this does not happen if `language = 'en'`. Convert them to periods if necessary:

```{r eval=FALSE}

data <- data %>%
  mutate(value = as.numeric(gsub(",", ".", value)))

```

- Time variables: Time information may be encoded differently depending on the table. Always check the metadata structure.

- Variable attributes: Long variable labels may contain additional information separated by commas. Clean these up based on your needs.

- Missing data: The databases use various symbols for missing or suppressed data (marked in `gen_signs()`). Handle these appropriately for your analysis.

## Summary: The Basic Workflow

1. Search: Use `gen_find()` or `gen_alternative_terms()` to discover relevant data
2. Explore: Check metadata with `gen_metadata_table()` to understand structure and filters
3. Filter: Identify the classification variables and category codes
4. Retrieve: Use `gen_table()` or `gen_cube()` with appropriate filtering parameters
5. Clean: Convert decimal separators, parse dates, clean variable labels
6. Analyse: Visualise, calculate statistics, or combine with other data sources
7. Monitor: Use `gen_modified_data()` to track updates

For more detailed information on additional API parameters available for filtering and customising your queries, see vignette("additional_parameter").

## Further Resources

- GENESIS API documentation: https://www-genesis.destatis.de/datenbank/online/docs/GENESIS-Webservices_Einfuehrung.pdf
- `{restatis}` package documentation: https://correlaid.github.io/restatis/
- German Federal Statistical Office: https://www.destatis.de/
